
---
output:
  html_document:
    self_contained: no
---

## The Cowboy Hat  

I've been a SAS user for about 30 years and when I was learning to use the 3D graphics procs in SAS I was introduced to a little program called the "Cowboy Hat". It creates a data set of values that when plotted look vaguely like a stylized 10 gallon cowboy hat. (sort of)

The code can be found here along with a simple graphic in SAS.

[Cowboy Hat](http://support.sas.com/documentation/cdl/en/hostwin/63285/HTML/default/viewer.htm#prodgraph.htm)

I've been taking the Data Science classes from Johns Hopkins on Coursera so I thought I'd try my hand at reproducing it here in R.


## The Data

Now lets run the code to create a data set of x, y and z values. I used for loops because that is what I know! I'm sure there are more than one way to accomplish this task....just like in SAS. The algorithm I used for coloring came from a comment on the blog Ripples by Andrew Wyer. Thanks Andrew!

[Ripples Post](https://aschinchon.wordpress.com/myself/#comments)

```{r}


library(rgl)
library(evd)


df <- data.frame(x=numeric(),
                 y=numeric(), 
                 z=numeric(), 
                 stringsAsFactors=FALSE) 

cnti <- 0


for (i in seq(-5, 5 ,by=.25) )
  
{ 
    
    for (j in  seq(-5 , 5,by=.25))
    {
      cnti <- cnti +1      
      z <- 4 * (sin(sqrt(i*i+j*j)))
      
      df[cnti,1] <- i
      df[cnti,2] <- j
      df[cnti,3] <- z
    }
}


cr <- abs(df$z)/max(abs(df$z))
cg <- abs(df$x)/max(abs(df$x))
cb <- abs(df$y)/max(abs(df$y))
```

## The Graphic

Now lets look at the graphic. You should be able to interact with it using your mouse scroll wheel and right click button.


```{r setup, results='asis'}
library(knitr)
knit_hooks$set(webgl = hook_webgl)
cat('<script type="text/javascript">', readLines(system.file('WebGL', 'CanvasMatrix.js', package = 'rgl')), '</script>', sep = '\n')
```




```{r testgl, webgl=TRUE}
# clear scene:

open3d()
# draw shperes in an rgl window
spheres3d(df$x, df$y, df$z, radius=0.05, color=rgb(cr,cg,cb), zlim=range(df$z))

```




## Some other possibilites

I also like using the tanh() function to replace the sin() function. When my wife was teaching and it was "Just say no to drugs" week at school I printed this one up and used the title "Beware the Black Hole of Drugs."

Try playing  around with some other trig functions and see what pops up!


